---
title: "GEDI PA Figures"
output: html_notebook
---

This notebook produces the figures for the Paper 'Global protected areas are effective for preserving forest structure and carbon'.

Set up environment
```{r}
#load any packages necessary
library(dplyr)
library(ggplot2)
local_fig_path <- '/gpfs/data1/duncansongp/GEDI_PA_paper/figures/'

#function for plotting up multiple plots with a single legend
grid_arrange_shared_legend <- function(..., ncol = length(list(...)), nrow = 1, position = c("bottom", "right")) {

  plots <- list(...)
  position <- match.arg(position)
  g <- ggplotGrob(plots[[1]] + theme(legend.position = position))$grobs
  legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
  lheight <- sum(legend$height)
  lwidth <- sum(legend$width)
  gl <- lapply(plots, function(x) x + theme(legend.position="none"))
  gl <- c(gl, ncol = ncol, nrow = nrow)

  combined <- switch(position,
                     "bottom" = arrangeGrob(do.call(arrangeGrob, gl),
                                            legend,
                                            ncol = 1,
                                            heights = unit.c(unit(1, "npc") - lheight, lheight)),
                     "right" = arrangeGrob(do.call(arrangeGrob, gl),
                                           legend,
                                           ncol = 2,
                                           widths = unit.c(unit(1, "npc") - lwidth, lwidth)))

  grid.newpage()
  grid.draw(combined)

  # return gtable invisibly
  invisible(combined)

}
```

Fig 1: Methods flowchart
```{r}
#Created Externally by V Leitold
```

Fig 2: Global GEDI height, RH50, cover & biomass
Amber to code - doesn't work directly at this point. 
```{r}
source("/gpfs/data1/duncansongp/amberliang/trends.Earth/git/GEDI_PA/fig2_script.R")
```

Fig 3: Global PAs for paper
```{r}
#Created externally by V Leitold
```

Fig 4: Number of GEDI shots per 1 km pixel
Amber to code so that this generates a figure directly from running notebook cell.
```{r}
#fig 4 a and 4b
#Coded  by M Liang
#GEDI shots distribution per 1km cell  Figure 4a
fpath <- "/gpfs/data1/duncansongp/GEDI_global_PA/WDPA_GEDI_extract3/"
dir(fpath)

shotInCell <- data.frame()
shotf=list.files(paste(fpath,"cell_stats",sep=""), full.names = TRUE)
# vars=c("x","y","gshot","status","pid","iso3")

for (d in shotf){
  print(d)
  cout <- read.csv(d)
  cout <- cout[!is.na(cout$pid),] 
  iso3 <- cout$country %>% unique() %>% na.omit()
  print(iso3)
  tmp <-  data.frame(iso3=cout$country, status=as.factor(cout$status), gshot=cout$gshot_counts, stringsAsFactors=F)
  shotInCell <- shotInCell %>% rbind(tmp, stringsAsFactors=F)
}

print(dim(shotInCell))

# plot
p <- shotInCell %>%
  ggplot( aes(x=gshot)) +
  geom_histogram( binwidth=5, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  labs(title = "GEDI shots in each matched 1km cell")+
       # ,subtitle = paste("Binwidth = 5; n =", format(nrow(shotInCell), big.mark=","), "; ","iso3 =", length(unique(shotInCell$iso3)), sep="")) +
  labs(x = "Number of GEDI shots in 1km Pixel")+theme_bw()

pzoom <- shotInCell %>%  #zoom in plot
  ggplot( aes(x=gshot)) +
  geom_histogram( binwidth=5, fill="#69b3a2", color="#e9ecef", alpha=0.9) +coord_cartesian(xlim=c(0,100))+
  labs(x = "Number of GEDI shots in 1km Pixel")+theme_bw()


#plot for shot in cell with inset map
p1I <- p + annotation_custom(ggplotGrob(pzoom), xmin = 280, xmax = 820, 
                       ymin = 250000, ymax =820000)+
  theme(text=element_text(family="Times", face="bold", size=14),
        legend.title=element_text(size=12), 
        legend.text=element_text( size=12))+
  labs(tag = "A")

ggplot2::ggsave(filename=paste('/gpfs/data1/duncansongp/GEDI_global_PA/figures/fig4a.png', sep=''), plot=p1I, width=6, height=6, units = "in", bg = "transparent")


#GEDI shots distribution per pa  Figure 4b
shotInPA <- data.frame()
shotpaf=list.files(paste(fpath, "pa_stats",sep=""), full.names = TRUE)

for (d in shotpaf){
  print(d)
  cout <- read.csv(d,header = TRUE)
  cout <- cout[!is.na(cout$pa_id),] 
  iso3 <- cout$iso3 %>% unique() %>% na.omit()
  print(iso3)
  tmp <-  data.frame(iso3=cout$iso3,pid=cout$pa_id, gshot=cout$count_1, stringsAsFactors=F)
  shotInPA <- shotInPA %>% rbind(tmp, stringsAsFactors=F)
}  
print(dim(shotInPA))
# plot
p2 <- shotInPA %>%
  ggplot( aes(x=gshot)) +
  geom_histogram( binwidth=500, fill="#FA8072", color="#FA8072", alpha=0.6) +
  labs(title = "GEDI shots in each matched PA")+
       # ,subtitle = paste("Binwidth = 500; ","n =", format(nrow(shotInPA), big.mark=","), "; iso3 =", length(unique(shotInPA$iso3)), sep=" ")) +
  theme_bw() +
  theme(
    plot.title = element_text(size=15)
  )+
  labs(x = "Number of GEDI shots in each PA")

p2zoom <- shotInPA %>%
  ggplot( aes(x=gshot)) +
  geom_histogram( binwidth=500, fill="#FA8072", color="#FA8072", alpha=0.6)+coord_cartesian(xlim=c(0,20000))+
  labs(x = "GEDI shots in each matched PA")+theme_bw()

#plot for shot in cell with inset map
p2I <- p2 + annotation_custom(ggplotGrob(p2zoom), ymin = 2800, ymax =10000, 
                      xmin = 450000, xmax =1400000)+
  theme(text=element_text(family="Times", face="bold", size=14),
        legend.title=element_text(size=12), 
        legend.text=element_text( size=12)) +
  labs(tag = "B")

ggplot2::ggsave(filename=paste('/gpfs/data1/duncansongp/GEDI_global_PA/figures/fig4b.png', sep=''), plot=p2I, width=6, height=6, units = "in", bg = "transparent")
```

Fig 5: Comparison of four metrics for globe
This should be overlapping histograms of each metric, potentially with breakdown by continent, PFT or biome in the topright of each figure to use up the space. 
Amber to code (fixing mess Laura made...)
```{r}
source("/gpfs/data1/duncansongp/amberliang/trends.Earth/git/GEDI_PA/fig5_script.R")

```

Fig 6: Comparison of four metrics for globe, divided by continent
Note: need to add whiskers for the SD 
```{r}
fpath <- "/Volumes/gsdata1/duncansongp/GEDI_global_PA/WDPA_GEDI_extract3/iso_full_nodup"
dir(fpath)

library(dplyr)
library(tidyr)
library(gridExtra)
library(cowplot)

#read in the full dataset
sub <-readRDS("/gpfs/data1/duncansongp/GEDI_global_PA/figures/all_iso_fullds.rds")

#change pa_status to text
status_factor <- rep('temp', nrow(metric_sub))
status_factor[which(metric_sub$pa_stat==0)] <- 'Control'
status_factor[which(metric_sub$pa_stat==1)] <- 'Protected Area'
status_factor <- as.factor(status_factor)
metric_sub <- cbind(metric_sub, status_factor)

#get the mean difference in the status groups by region.
#AGBD by Continent
sub$region <- fct_explicit_na(sub$region)
sub %>%
  group_by(status, region) %>%
  dplyr::summarise(mean = mean(AGBD, na.rm=TRUE)) -> regional_means_AGBD

mean_control <- regional_means_AGBD[which(regional_means_AGBD$status==0),]
mean_treatment <- regional_means_AGBD[which(regional_means_AGBD$status==1),]
mean_diffs <- 100*((mean_treatment$mean - mean_control$mean) / mean_treatment$mean)

mp <- data.frame(mean_diffs=mean_diffs[-length(mean_diffs)], continent=mean_control$region[-length(mean_diffs)])

mp$colour <- ifelse(mp$mean_diffs < 0, "firebrick1","steelblue")
mp$hjust <- ifelse(mp$mean_diffs > 0, 1.3, -0.3)

agbd_continent <- ggplot(mp, aes(y=mean_diffs, x=continent, hjust = hjust)) + 
  geom_bar(stat = "identity", aes(fill = colour)) +
  labs(y = 'Mean AGBD Difference (% Mg/ha)') +
  labs(x = 'Continent') +
  theme(legend.position = 'none', text = element_text(size=40))

#Height by Continent
sub$region <- fct_explicit_na(sub$region)
sub %>%
  group_by(status, region) %>%
  dplyr::summarise(mean = mean(rh_098, na.rm=TRUE)) -> regional_means_ht

mean_control <- regional_means_ht[which(regional_means_ht$status==0),]
mean_treatment <- regional_means_ht[which(regional_means_ht$status==1),]
mean_diffs <- mean_treatment$mean - mean_control$mean

mp <- data.frame(mean_diffs=mean_diffs[-length(mean_diffs)], continent=mean_control$region[-length(mean_diffs)])

mp$colour <- ifelse(mp$mean_diffs < 0, "firebrick1","steelblue")
mp$hjust <- ifelse(mp$mean_diffs > 0, 1.3, -0.3)

ht_continent <- ggplot(mp, aes(y=mean_diffs, x=continent, hjust = hjust)) + 
  geom_bar(stat = "identity") +
  labs(y = 'Mean Height Difference (m)') +
  labs(x = 'Continent') +
  theme(legend.position = 'none', text = element_text(size=40))

#PAI by Continent
sub$region <- fct_explicit_na(sub$region)
sub %>%
  group_by(status, region) %>%
  dplyr::summarise(mean = mean(pai, na.rm=TRUE)) -> regional_means_pai

mean_control <- regional_means_pai[which(regional_means_pai$status==0),]
mean_treatment <- regional_means_pai[which(regional_means_pai$status==1),]
mean_diffs <- 100*(mean_treatment$mean - mean_control$mean) / mean_treatment$mean

mp <- data.frame(mean_diffs=mean_diffs[-length(mean_diffs)], continent=mean_control$region[-length(mean_diffs)])

mp$colour <- ifelse(mp$mean_diffs < 0, "steelblue","firebrickred1")
mp$hjust <- ifelse(mp$mean_diffs > 0, 1.3, -0.3)

pai_continent <- ggplot(mp, aes(y=mean_diffs, x=continent, hjust = hjust)) + 
  geom_bar(stat = "identity") +
  labs(y = 'Mean PAI Difference (%)') +
  labs(x = 'Continent') +
  theme(legend.position = 'none', text = element_text(size=40))

#Cover by Continent
sub$region <- fct_explicit_na(sub$region)
sub %>%
  group_by(status, region) %>%
  dplyr::summarise(mean = mean(cover, na.rm=TRUE)) -> regional_means_cover

mean_control <- regional_means_cover[which(regional_means_cover$status==0),]
mean_treatment <- regional_means_cover[which(regional_means_cover$status==1),]
mean_diffs <- 100*(mean_treatment$mean - mean_control$mean) / mean_treatment$mean

mp <- data.frame(mean_diffs=mean_diffs[-length(mean_diffs)], continent=mean_control$region[-length(mean_diffs)])

mp$colour <- ifelse(mp$mean_diffs < 0, "firebrickred1","steelblue")
mp$hjust <- ifelse(mp$mean_diffs > 0, 1.3, -0.3)

cover_continent <- ggplot(mp, aes(y=mean_diffs, x=continent, hjust = hjust)) + 
  geom_bar(stat = "identity") +
  labs(y = 'Mean Cover Difference (%)') +
  labs(x = 'Continent') +
  theme(legend.position = 'none', text = element_text(size=40))

png_file<-paste(filename="/gpfs/data1/duncansongp/GEDI_global_PA/figures/fig6.png")
png(width=1800,height=2000,file=png_file)
grid.arrange(ht_continent, cover_continent, pai_continent, agbd_continent, ncol=2)
dev.off()


```

Fig 7: Map of globe with barchart of mean difference (m for height, %diff for PAI, Cover, AGBD) placed in each corresponding continent. Vero to take first crack. Code for Fig 6 can be modified to make this. 
SD plotted as whiskers for each bar. 
```{r}

```

Fig 8: Comparison of four metrics for different biome classes (same as Fig 6 but for biome)
```{r}

```


Fig 9: Comparison of four metrics for the top 10 countries with most PAs (or similar)
```{r}

```

Fig 10: Distributions of biomass within protected areas vs. outside of protected areas by biome, urban mask applied
```{r}

```

Fig 11: Country-level PA distribution figure (e.g. Brazil)
```{r}
#Created externally by V Leitold
```

Fig 12: Relative importance of variables to determine biomass preservation efficacy 
```{r}

```

Tables and final statistical moments aggregation
```{r}
source("/gpfs/data1/duncansongp/amberliang/trends.Earth/git/GEDI_PA/final_stats_agg.R")

```


