---
title: "GEDI PA Figures"
output: html_notebook
---

This notebook produces the figures for the Paper 'Global protected areas are effective for preserving forest structure and carbon'.

Fig 1: Methods flowchart
```{r}
#Created Externally by V Leitold
```

Fig 2: Global GEDI height, RH50, cover & biomass
```{r}
#Placeholder exists, but we will create our own version with the final data used 
```

Fig 3: Methods workflow for paper
```{r}
#Created externally by V Leitold
```

Fig 4: Number of GEDI shots per 1 km pixel
```{r}
#Created externally by M Liang
```

Fig 5: Comparison of four metrics for globe
```{r}

```

Fig 6: Comparison of four metrics for globe, divided by continent
```{r}

```

Fig 7: Comparison of four metrics for the top 10 countries with most PAs (or similar)
```{r}

```

Fig 8: Comparison of four metrics for different biome classes
```{r}

```

Fig 9: Country-level PA distribution figure (e.g. Brazil)
```{r}
#Created externally by V Leitold
```

Fig 10: Relative importance of variables to determine biomass preservation efficacy 
```{r}

```




